### 2016-01-07 p2/comparisons.tex
Finished conclusion

We consider the reconciliation algorithm described here an improvement over
the one derived during our previous research (\cite{NREC})
as the previous algorithm not only fails to propagate all possible commands
wherever it is possible (that is, where it does not break the filesystem),
but the current algorithm is also simpler.
This is because the previous reconciliation algorithm excludes
commands from being propagated which must be preceded by a command that conflicts.

@@

As a test of the proposed reconciliation algorithm, 
we compare our definition of conflicting
updates to how conflicts are defined by Balasubramaniam and Pierce
in their state-based approach implemented in the Unison synchronizer (\cite{BP}).
As we noted in \cite{NREC}, the update detector they describe provides a safe estimate of nodes
(paths) at which updates occurred.
It marks some nodes as \emph{dirty} in a way that we know that at non-dirty nodes
the filesystems (replicas) have not changed between their common original state and current state
(see Definition 3.1.1 in \cite{BP}).
The \emph{dirty} marks are also up-closed, that is, all ancestor nodes of a dirty node
are also dirty (Fact 3.1.3 in \cite{BP}).
And finally, in our notation, 
a conflict is detected between replicas $\FS_A$ and $\FS_B$ at node $n$
if $n$ is marked as dirty in both $\FS_A$ and $\FS_B$, and
$\FS_A(n)\neq\FS_B(n)$, and $n$ does not point to a directory in both replicas.
(In other words, 
$\FS_A(n)\not\in\setd \vee \FS_B(n)\not\in\setd$; see section 4.1 in \cite{BP}.
Let us note that there is an alternative approach to defining an algorithm for
the same synchronizer by Pierce and Vouillon in \cite{PV}.)

>>>>

We consider the reconciliation algorithm described here an improvement over
the one derived during our previous research \cite{NREC}
as the previous algorithm not only fails to propagate all possible commands
wherever it is possible (that is, where it does not break the filesystem),
but the current algorithm is also simpler.
This is because the previous reconciliation algorithm excludes
commands from being propagated which must be preceded by a command that conflicts.

@@

As a test of the proposed reconciliation algorithm, 
we compare our definition of conflicting
updates to how conflicts are defined by Balasubramaniam and Pierce
in their state-based approach implemented in the Unison synchronizer \cite{BP}.
As we noted in \cite{NREC}, the update detector they describe provides a safe estimate of nodes
(paths) at which updates occurred.
It marks some nodes as \emph{dirty} in a way that we know that at non-dirty nodes
the filesystems (replicas) have not changed between their common original state and current state
(see Definition 3.1.1 in \cite{BP}).
The \emph{dirty} marks are also up-closed, that is, all ancestor nodes of a dirty node
are also dirty (Fact 3.1.3 in \cite{BP}).
And finally, in our notation, 
a conflict is detected between replicas $\FS_A$ and $\FS_B$ at node $n$
if $n$ is marked as dirty in both $\FS_A$ and $\FS_B$, and
$\FS_A(n)\neq\FS_B(n)$, and $n$ does not point to a directory in both replicas.
(In other words, 
$\FS_A(n)\not\in\setd \vee \FS_B(n)\not\in\setd$; see section 4.1 in \cite{BP}.
Let us note that there is an alternative approach to defining an algorithm for
the same synchronizer by Pierce and Vouillon in \cite{PV}.)

### 2016-01-07 p2/conclusion.tex
New reference
New references and intro
Finished conclusion

% TODO
... ... ...

(Same approach to other systems)
\medskip

% Further research

Apart from constructing an algebra,
there are many other ways in which further work can extend the current results.
We think one of the most interesting extensions would be to
consider reconciling not only two, but more replicas in a single step and
prove the correctness and maximality of the algorithm proposed.
A related problem is to model cases where reconciliation cannot
complete fully
or if only a subset of the replicas are reconciled 
(e.g. due to network partitioning),
both of which would result in a state where different replicas
have different common ancestors, that is,
the updates specific to the replicas start from different points
in the update history of the filesystem.
Parker et al. (\cite{PPRS}) and Cox and Josephson (\cite{CJ})
describe version vectors (update histories) kept as metadata that could solve this problem.
Extending the algebraic approach to represent individual updates to files
in their modification histories (as described in \cite{CJ})
as separate commands could also enable it to reconcile otherwise
conflicting updates and resolve partial reconciliations.

An extension of these results could also entail relaxing
the simplifying assumptions we used.
Identifying which results depend on the assumption that $|\setd|=1$
and devising updated proofs or algorithms,
or a pre-processing step that reconciles meta-information in directories
can enable the algebraic system or an implementation
to identify and resolve a wider range of conflicts.
Similarly, designing pre- and post-processors that
present certain changes not as removal and creation of a node,
but as a rename,
could contribute to the user-friendliness of the reconciler
and help avoid human error.

>>>>

Apart from constructing an algebra,
there are many other ways in which further work can extend the current results.
An important extension would be to
consider reconciling not only two, but more replicas in a single step and
prove the correctness and maximality of the algorithm proposed.

A related problem is to extend the system and the proofs
to allow for cases where reconciliation cannot
complete fully
or if only a subset of the replicas are reconciled 
(e.g. due to network partitioning),
both of which would result in a state where different replicas
have different common ancestors, that is,
the updates specific to the replicas start from different points
in the update history of the filesystem.
Existing research can offer pointers as to how such cases can be modelled
in our algebraic system.
Parker et al. \cite{PPRS} and Cox and Josephson \cite{CJ}
describe version vectors (update histories) kept as metadata,
while Chong and Hamadi present distributed algorithms that allow incremental synchronization \cite{CH}.
Representing individual updates to files
in their modification histories (as described in \cite{CJ})
as separate commands could also enable an algebraic synchronizer to reconcile otherwise
conflicting updates and resolve partial reconciliations.

An extension of our results could also entail relaxing
the simplifying assumptions we used.
Identifying which results depend on the assumption that $|\setd|=1$
and devising updated proofs and algorithms,
or specifying a pre-processing step that reconciles meta-information in directories
can enable the synchronizer
to identify and resolve a wider range of conflicts.
Similarly, designing pre- and post-processors that
present removing and creating a node
not as separate updates but as a single rename (move),
could contribute to the user-friendliness of the reconciler
and help avoid human error.

And finally, we hope that this work, together with \cite{NREC}, provides
a blueprint of constructing an algebra of commands for different storage protocols
(e.g. XML trees, mailbox folders, generic relational databases, etc.),
and of demonstrating the adequacy and completeness of update and conflict detection and reconciliation
algorithms defined over it.
This, in turn, can offer formal verification of the algorithms underlying
specific implementations in a variety of synchronizers.
Alternatively, by generalizing the parent--child relationships between filesystem nodes,
the demonstrated properties of minimal sequences of commands
and conditional operation of sequences ($\workssign$)
may also contribute to future research into algebraic structures
constrained by predefined sparse connections between their elements.

### 2016-01-07 p2/def-commands.tex
Style updates
Finished conclusion

Let us consider what kind of information is usually encoded in filesystem operations.
A minimal set of commands, based on the most frequent tools implemented by filesystems,
may be $create(n,\valv)$, which creates a file or directory ($\valv$) at the empty node $n$;
$edit(n,\valv)$, which replaces the earlier value at $n$ with $\valv$;
and $remove(n)$, which deletes the value at $n$---where 
$n\in\setn$ and $\valv\in\setv$ (but $\valv\neq\empt$).
Regarding their output, that is, the state of the filesystem at $n$
after applying the command,
we know that after $create$ or $edit$, $\FS(n)\neq\empt$, whereas after $remove$,
$\FS(n)$ will be $\empt$. 
However, from \cite{NREC} and \cite{CBNR} we know that a useful set of axioms
will in some cases need to distinguish between, for example,
$edit$s that result in directories ($edit(n,\vald)$) and
ones that result in files ($edit(n,\valf)$), and treat them as separate commands,
as their behaviors are quite different when combined with other commands.
Indeed, Bill Zissimopoulos' work
demonstrated (\cite{BZ})
that extending this distinction to more commands ultimately simplifies
the definition of conflicting commands, as our model will then able to predict the behavior of commands
more precisely.
In other words, encoding the type of the output ($\cchard$, $\ccharf$ or $\ccharb$) in the commands is definitely useful.
At the same time, there is never any need to consider the
exact output \emph{value} of a command,
as the success or failure of filesystem commands only depends on the types of values in the filesystem.

@@

\begin{mydef}[Effect of commands]
We use a \emph{replacement operator} of the form
$\FS[\valv/n]$ to denote a filesystem derived from $\FS$:
\[ \FS[\valv/n](m) :=
   \begin{cases}
   \valv &\mbox{if~} m=n\\
   \FS(m) &\mbox{otherwise.}
   \end{cases}
\]
The effect of the commands is then as follows:
\begin{align*}
&\cbrk\aFS = \fsbroken \\
&\cxynv\aFS = 
   \begin{cases}
   \fsbroken &\mbox{if~} \FS=\fsbroken\\
   \fsbroken &\mbox{if~} \FS(n)\not\in\setvx{X}\\
   \fsbroken &\mbox{if~} \FS[\valvy/n] \mbox{~violates the tree property}\\
   \FS[\valvy/n] &\mbox{otherwise.}
   \end{cases}
\end{align*}
\end{mydef}

>>>>

Let us consider what kind of information is usually encoded in filesystem operations.
A minimal set of commands, based on the most frequent tools implemented by filesystems,
may be $create(n,\valv)$, which creates a file or directory ($\valv$) at the empty node $n$;
$edit(n,\valv)$, which replaces the earlier value at $n$ with $\valv$;
and $remove(n)$, which deletes the value at $n$---where 
$n\in\setn$ and $\valv\in\setv$ (but $\valv\neq\empt$).
Regarding their output, that is, the state of the filesystem at $n$
after applying the command,
we know that after $create$ or $edit$, $\FS(n)\neq\empt$, whereas after $remove$,
$\FS(n)$ will be $\empt$. 
However, from \cite{NREC} and \cite{CBNR} we know that a useful set of axioms
will in some cases need to distinguish between, for example,
$edit$s that result in directories ($edit(n,\vald)$) and
ones that result in files ($edit(n,\valf)$), and treat them as separate commands,
as their behaviors are quite different when combined with other commands.
Indeed, Bill Zissimopoulos' work
demonstrated \cite{BZ}
that extending this distinction to more commands ultimately simplifies
the definition of conflicting commands, as our model will then able to predict the behavior of commands
more precisely.
In other words, encoding the type of the output ($\cchard$, $\ccharf$ or $\ccharb$) in the commands is definitely useful.
At the same time, there is never any need to consider the
exact output \emph{value} of a command,
as the success or failure of filesystem commands only depends on the types of values in the filesystem.

@@

\begin{mydef}[Effect of commands]
We use a \emph{replacement operator} of the form
$\FS[\valv/n]$ to denote a filesystem derived from $\FS$:
\[ \FS[\valv/n](m) =
   \begin{cases}
   \valv &\mbox{if~} m=n\\
   \FS(m) &\mbox{otherwise.}
   \end{cases}
\]
The effect of the commands is then as follows:
\begin{align*}
&\cbrk\aFS = \fsbroken \\
&\cxynv\aFS = 
   \begin{cases}
   \fsbroken &\mbox{if~} \FS=\fsbroken\\
   \fsbroken &\mbox{if~} \FS(n)\not\in\setvx{X}\\
   \fsbroken &\mbox{if~} \FS[\valvy/n] \mbox{~violates the tree property}\\
   \FS[\valvy/n] &\mbox{otherwise.}
   \end{cases}
\end{align*}
\end{mydef}

### 2016-01-07 p2/def-filesystem.tex
Style updates
Minor update
Finished conclusion

Every working filesystem has a so-called \emph{tree property}, which means that
if the filesystem is not empty at a node, and the node has a parent,
then there must be a directory at the parent node.
To model this, in $\setv$ we select another special value
to represent directories, as we assume that apart from their location
in the filesystem, all directories are equal.
We do this because, as Bill Zissimopoulos pointed 
out (\cite{BZ}),
we often do not want to consider metadata stored in
directories (e.g. permission settings) during synchronization,
as these are not generally understood well by users,
and, if needed, conflict resolution on these settings can be easily automated.
We will also see that this assumption makes it possible
to define a maximal reconciliation algorithm.

@@

In this paper, $n$, $m$ and $o$ denote nodes in $\setn$.
We use $\valf$ for an arbitrary element in $\setf$, 
and by $\valvx$ we mean a value of type $X$ in $\setvx{X}$.

>>>>

Every working filesystem has a so-called \emph{tree property}, which means that
if the filesystem is not empty at a node, and the node has a parent,
then there must be a directory at the parent node.
To model this, in $\setv$ we select another special value
to represent directories, as we assume that apart from their location
in the filesystem, all directories are equal.
We do this because, as Bill Zissimopoulos pointed 
out \cite{BZ},
we often do not want to consider metadata stored in
directories (e.g. permission settings) during synchronization,
as these are not generally understood well by users,
and, if needed, conflict resolution on these settings can be easily automated.
We will also see that this assumption makes it possible
to define a maximal reconciliation algorithm.

@@

In this paper, $\FS$ and $\GS$ denote filesystems,
and $n$, $m$ and $o$ are nodes in $\setn$.
We use $\valf$ for an arbitrary element in $\setf$, 
and by $\valvx$ we mean a value of type $X$ in $\setvx{X}$.

### 2016-01-07 p2/intro.tex
Finished conclusion

In this paper we investigate a command-based approach
to synchronizing filesystem trees.
Presented with multiple copies (replicas) of the same filesystem that have been independently modified,
we regard the aim of the synchronizer to modify these replicas further
to make them as similar as possible.
The synchronization algorithm we describe
follows the two main steps described by Balasubramaniam and Pierce (\cite{BP}):
update detection, where we identify modifications that have been applied to the replicas
since they diverged and represent these as sequences of commands,
and reconciliation, where we identify commands that can be propagated to other replicas and do so.
The remaining commands, which could not be propagated, represent conflicting updates;
resolving these requires a separate system or human intervention.

In this paper we build on previous work done 
by Ramsey and Csirmaz
on an algebraic approach to synchronization
(\cite{NREC}),
and add to the theoretical understanding
of synchronization by providing rigorous proofs that the algorithms
we describe work as intended
and offer a solution in our framework that cannot be improved on.

@@

In terms of the usual classification of synchronizers (\cite{TSR, PV, SSH}),
our approach to reconciliation is therefore operation-based as we investigate sets of commands,
while the update detection algorithm makes the full synchronizer 
similar to state-based synchronizers 
as its starting point is the states of the replicas themselves.
Moreover, our ordering algorithms only depend on the properties of commands 
such as commutativity or idempotence,
and therefore they can be classified as semantic schedulers.
This is in opposition to syntactic scheduling, which uses 
the origin and original temporal order of updates, 
and which may give rise to otherwise resolvable conflicts (\cite{SSH}).
Ordering plays a central role in other operation-based synchronizers.
We refer to a survey by Saito and Shapiro of so-called optimistic replication algorithms (\cite{SSH});
or, in particular,
to IceCube, where multiple orders are tested to find an acceptable one (\cite{KRSD}, see also \cite{MPV}),
or Bayou, where reconciling updates happens by redoing them in a globally determined order \cite{TTPDSH}.

>>>>

In this paper we investigate a command-based approach
to synchronizing filesystem trees.
Presented with multiple copies (replicas) of the same filesystem that have been independently modified,
we regard the aim of the synchronizer to modify these replicas further
to make them as similar as possible.
The synchronization algorithm we describe
follows the two main steps described by Balasubramaniam and Pierce \cite{BP}:
update detection, where we identify modifications that have been applied to the replicas
since they diverged and represent these as sequences of commands,
and reconciliation, where we identify commands that can be propagated to other replicas and do so.
The remaining commands, which could not be propagated, represent conflicting updates;
resolving these requires a separate system or human intervention.

In this paper we build on our previous work done 
with Prof. Norman Ramsey
on an algebraic approach to synchronization
\cite{NREC},
and add to the theoretical understanding
of synchronization by providing rigorous proofs that the algorithms
we describe work as intended
and offer a solution in our framework that cannot be improved on.

@@

In terms of the usual classification of synchronizers \cite{TSR, PV, SSH},
our approach to reconciliation is therefore operation-based as we investigate sets of commands,
while the update detection algorithm makes the full synchronizer 
similar to state-based synchronizers 
as its starting point is the states of the replicas themselves.
Moreover, our ordering algorithms only depend on properties of commands 
such as commutativity or idempotence,
and therefore they can be classified as semantic schedulers.
This is in opposition to syntactic scheduling, which uses 
the origin and original temporal order of updates, 
and which may give rise to otherwise resolvable conflicts \cite{SSH}.
In general, ordering plays a central role in operation-based synchronizers.
We refer to a survey by Saito and Shapiro of so-called optimistic replication algorithms \cite{SSH};
and, in particular,
to IceCube, where multiple orders are tested to find an acceptable one (\cite{KRSD}, see also \cite{MPV}),
or Bayou, where reconciling updates happens by redoing them in a globally determined order \cite{TTPDSH}.

### 2016-01-07 p2/ordering.tex
Renamed directories
Finished conclusion

\medskip

We can now define our algorithm to generate orders of sets of commands.

@@

\begin{mylem}\label{simple_reorder_equiv}
For all simple sequences $T$ and $S\nequiv\cbrk$,
$T\equiv S \Leftrightarrow T\in\orderset{S}$.
\end{mylem}
\begin{proof}
First, as $S$ only has at most one command on each node,
any sequence in $\orderset{S}$ either breaks a filesystem,
or it leaves it in the same state as $S$.

>>>>

\bigskip

\noindent
We can now define our algorithm to generate orders of sets of commands.

@@

\begin{mylem}\label{simple_reorder_equiv}
For all simple sequences $T$ and $S\nequiv\cbrk$,
\[ T\equiv S \Longleftrightarrow T\in\orderset{S}. \]
\end{mylem}
\begin{proof}
First, as $S$ only has at most one command on each node,
any sequence in $\orderset{S}$ either breaks a filesystem,
or it leaves it in the same state as $S$.
